\section{Introduction }

The previous chapter described the AI-Scientist architecture as an orchestrated verification pipeline. This chapter
maps that design onto the actual implementation. The goal is not to list every file mechanically, but to explain
how the main modules cooperate to realize evidence-grounded, self-critical, traceable verification in code, in the
spirit of agentic and retrieval-augmented systems such as AutoGen, RAG, and REALM \cite{wu2023autogen,lewis2020rag,guu2020realm}. In
other words, this chapter shows how the architectural ideas become a working software system that can be executed,
inspected, and discussed as part of a thesis.

The implementation choices also echo the broader design patterns seen in recent retrieval-first and self-refining
systems, where structured orchestration, explicit evidence flow, and optional model refinement are used to improve
trustworthiness without sacrificing reproducibility \cite{shinn2023reflexion,yao2023treeofthoughts,asai2024selfrag}.

The implementation is organized as a single Python package, \texttt{ai\_scientist}, with the agents in an
\texttt{agents} subpackage and a set of supporting modules for configuration, data models, corpus access,
ingestion, retrieval utilities, scope control, and the optional language-model layer. A Streamlit user interface
and a programmatic API sit on top of the same package. This separation is intentional: the package contains the
reasoning logic, while the interfaces are thin clients that drive it.

\begin{figure}[!htbp]
\centering
\begin{tikzpicture}[
  scale=0.96,
  transform shape,
  font=\small,
  layer/.style={draw, rounded corners, align=center, text width=34mm, minimum height=11mm, minimum width=34mm, fill=blue!15},
  core/.style={draw, rounded corners, align=center, text width=56mm, minimum height=15mm, minimum width=56mm, fill=yellow!28},
  resultbox/.style={draw, rounded corners, align=center, text width=35mm, minimum height=11mm, minimum width=35mm, fill=orange!18},
  arrow/.style={-{Stealth[round]}, thick}
]
\node[layer] (ui) {Interfaces\\ \scriptsize Streamlit + FastAPI};
\node[core, below=8mm of ui] (orch) {Orchestrator\\ \scriptsize workflow state + trace};
\node[layer, below left=12mm and 12mm of orch] (agents) {Agent Modules\\ \scriptsize retrieval, verification, critique};
\node[layer, below right=12mm and 12mm of orch] (coremods) {Core Modules\\ \scriptsize models, config, corpus, scope};
\node[resultbox, below=12mm of orch] (results) {Outputs\\ \scriptsize verified claims, summary, report};

\draw[arrow] (ui) -- (orch);
\draw[arrow] (orch) -- (agents);
\draw[arrow] (orch) -- (coremods);
\draw[arrow] (agents) -- (results);
\draw[arrow] (coremods) -- (results);
\end{tikzpicture}
\caption{Implementation stack of the AI-Scientist package}
\label{fig:implementation_stack}
\end{figure}

\section{Implementation Structure }

At the highest level, the codebase is divided into:

\begin{enumerate}
\item \textbf{agent modules} under \path{ai_scientist/agents/}, which implement retrieval, claim extraction,
verification, critique, synthesis, reporting, and live retrieval;
\item \textbf{the orchestrator} in \path{ai_scientist/agents/orchestrator.py}, which wires the agents into the
end-to-end workflow;
\item \textbf{core modules} --- \path{models.py}, \path{config.py}, \path{corpus.py}, \path{ingestion.py},
\path{utils.py}, \path{scope.py}, and \path{llm.py} --- which provide the data types, settings, evidence access,
and shared logic;
\item \textbf{baselines} in \path{baselines.py}, which implement the single-agent and RAG comparison systems on
top of the same retrieval and scoring code; and
\item \textbf{interfaces}: a Streamlit application (\path{app.py}) and a FastAPI service
(\path{ai_scientist/api.py}).
\end{enumerate}

This layered structure keeps the reasoning system reusable. The same analyze question(...) entry point is used by the user
interface, the API, and the baselines harness, so that what is
demonstrated is the real system rather than a simplified mock-up. That
alignment also makes it easier to compare the implementation with
retrieval-oriented benchmarks such as BEIR and KILT \cite{thakur2021beir,petroni2020kilt,ragsurvey2024}.

This also makes the implementation easier to explain in the thesis. A
single call path supports interactive use, programmatic access, and
benchmark comparison, which keeps the narrative compact and the software
story coherent, while preserving the kind of transparent control surface
often emphasized in retrieval-first systems \cite{karpukhin2020dpr,khattab2020colbert,lewis2020rag}.

\begin{table}[H]
\centering
\small
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.15}
\caption{Implementation layers and their purpose}
\begin{tabularx}{0.98\linewidth}{>{\raggedright\arraybackslash}p{0.24\linewidth} >{\raggedright\arraybackslash}X}
\toprule
Layer & Purpose \\
\midrule
Interfaces & Provide user interaction through Streamlit and API access through FastAPI \\
Orchestrator & Coordinates the pipeline, manages traceability, and handles early exits \\
Agent modules & Perform retrieval, claim extraction, verification, critique, synthesis, and reporting \\
Core modules & Store shared data structures, configuration, scope logic, and corpus access \\
Baselines & Offer comparison systems on the same retrieval substrate \\
\bottomrule
\end{tabularx}
\end{table}

\begin{figure}[!htbp]
\centering
\begin{tikzpicture}[
  scale=0.92,
  transform shape,
  font=\small,
  paper/.style={draw, rounded corners, align=center, minimum height=10mm, minimum width=31mm, fill=blue!15},
  claim/.style={draw, rounded corners, align=center, minimum height=10mm, minimum width=31mm, fill=blue!15},
  evidence/.style={draw, rounded corners, align=center, minimum height=10mm, minimum width=31mm, fill=blue!15},
  assess/.style={draw, rounded corners, align=center, minimum height=10mm, minimum width=33mm, fill=yellow!25},
  report/.style={draw, rounded corners, align=center, minimum height=10mm, minimum width=31mm, fill=orange!20},
  arrow/.style={-{Stealth[round]}, thick}
]
\node[paper] (paper) {PaperDocument};
\node[claim, right=10mm of paper] (claim) {StructuredClaim};
\node[evidence, right=10mm of claim] (evi) {EvidenceSnippet};
\node[assess, below=10mm of claim] (assess) {ClaimAssessment};
\node[report, below=10mm of assess] (report) {FinalReport};

\draw[arrow] (paper) -- (claim);
\draw[arrow] (claim) -- (evi);
\draw[arrow] (claim) -- (assess);
\draw[arrow] (evi) -- (assess);
\draw[arrow] (assess) -- (report);
\end{tikzpicture}
\caption{Typed data flow used inside the implementation}
\label{fig:typed_data_flow}
\end{figure}

\section{Typed Data Models }

The pipeline communicates through explicit dataclasses defined in \path{models.py}. A \texttt{PaperDocument}
carries a paper's identifier, title, abstract, year, authors, and topics. A \texttt{StructuredClaim} records a
claim's text, type, source paper, source sentence, and focus terms. An \texttt{EvidenceSnippet} pairs a sentence
with its overlap score and source. A \texttt{ClaimAssessment} bundles a claim with its verdict, confidence,
evidence list, and rationale. Finally, a \texttt{FinalReport} aggregates the question, summary, verified claims,
critique notes, retrieved papers, rendered Markdown, and the iteration trace. This typed flow is consistent with
the structured claim and evidence representations used in factuality-oriented verification work \cite{thorne2018fever,wadden2020scifact,min2023factscore}.

These typed structures are important because every stage writes to a shared, serializable interface rather than
passing unstructured text. This is what makes the system auditable: the report is not a paragraph that happens to
mention sources, but a structured object in which each conclusion is explicitly linked to its evidence.

The dataclass layer also provides a stable vocabulary for the rest of the thesis. Once papers, claims, evidence,
and reports are named explicitly in code, the later evaluation and discussion chapters can refer to them
consistently without ambiguity.

\section{The Orchestrator }

The practical center of the implementation is \texttt{MultiAgentResearchSystem}
in \path{orchestrator.py}. Its \texttt{analyze\_question} method
realizes the workflow described in Chapter~4: it applies the scope
guard, retrieves evidence, extracts claims, verifies them through a
refinement routine, runs the Critic, and optionally re-verifies weak
claims before synthesizing the summary. The method also handles the
early exits used by the system: if the question is out of scope, if
retrieval returns no papers, or if no aligned claims can be extracted,
it returns a report that explains the situation and preserves
traceability, which is in line with the controlled orchestration patterns used in modern multi-agent systems \cite{wu2023autogen,shinn2023reflexion}.

Adaptive re-verification is implemented in the private
\texttt{\_refine\_assessments} and \texttt{\_run\_second\_pass}
routines. After the first pass, the Critic's identify reverification targets method produces follow-up
decisions; for each one, the Orchestrator retrieves additional papers
for that claim, merges them with the first-pass evidence, and
re-verifies the claim with contradiction resolution preferred. Every
step appends a human-readable line to the iteration trace, which is
what the user ultimately sees as the orchestration trace.

This implementation choice keeps the adaptive loop transparent. Instead of hiding refinement inside a silent
retry, the system exposes the extra pass as part of the report, which makes both debugging and thesis writing
more straightforward, and it aligns well with the critique-and-refine style documented in adaptive reasoning work \cite{yao2023treeofthoughts,asai2024selfrag}.

\section{Retrieval and the Layered Evidence Pool }

The Retrieval agent in \path{agents/retrieval.py} implements the layered evidence model. Its \texttt{\_local\_pool}
builds a deduplicated pool that lists uploaded papers before curated-corpus papers, and \texttt{\_local\_rank\_score}
adds the source-dependent bonus so that uploads outrank corpus papers at comparable relevance. The
\texttt{\_rank\_subset} routine ranks candidates by this source-weighted score but gates admission on the raw
query-coverage clearing the configured minimum, so the bonus can reorder evidence but never inject an irrelevant
paper, which is consistent with evidence-ranking practice in dense retrieval systems \cite{karpukhin2020dpr,khattab2020colbert,thakur2021beir}.

When live retrieval is enabled, \texttt{retrieve} layers API results on top of the local pool. Live papers pass
through a quality filter that rewards adequate abstract length, query relevance in title and abstract, recency, and
the presence of research-indicator terms, and they are merged with a strict source priority: uploads first, then
corpus, then live. A title-similarity check removes near-duplicates so the same paper is not counted twice across
sources. The agent is wrapped so that any network failure falls back cleanly to the local pool, keeping the system
usable offline.

The retrieval layer therefore serves both ranking and provenance. It lets the system prefer the user's own
materials, preserve a dependable offline core, and still extend outward when live evidence adds value.
That layered design mirrors the broader retrieval-augmented generation literature, where grounding improves
answer quality precisely because evidence selection remains explicit \cite{lewis2020rag,guu2020realm,ragsurvey2024}.

\section{Retrieval Utilities }

The relevance measures live in \path{utils.py}. Tokenization normalizes case, strips stopwords, and applies a
small morphological normalization so that, for example, plural and singular forms of common research terms match.
On top of tokenization, \texttt{overlap\_score} computes Jaccard overlap for short-text comparison, and
\texttt{coverage\_score} computes the length-robust query coverage used for paper-level ranking. Keeping both
measures in one module makes the design decision from Chapter~3 --- coverage for ranking, Jaccard for
sentence-level evidence --- explicit and reusable across the verifier, the retriever, and the baselines.

Because these utilities are shared across the pipeline, the implementation stays internally consistent. The same
similarity intuition is used when ranking papers, checking claim support, and comparing the baselines, which
reduces the chance of mismatched behaviour between components, a useful property in systems that combine retrieval
and factuality evaluation \cite{min2023factscore,es2023ragas}.

\section{Claim Extraction }

The Claim Extraction agent in \path{agents/claim_extraction.py} scans each retrieved paper sentence by sentence. A
sentence becomes a candidate claim if it contains a positive research cue or shares content tokens with the
question. Each candidate is wrapped in a \texttt{StructuredClaim} whose type is inferred from surface cues ---
limitation, performance, finding, implication, or observation --- and whose focus terms are the intersection of the
question and sentence tokens. Extraction stops once a configured maximum number of claims is reached, keeping the
workflow focused on the statements most relevant to the query, similar in spirit to claim-centric factuality
pipelines that prefer compact evidence units over long free-form passages \cite{thorne2018fever,wadden2020scifact,honovich2022true}.

The bounded extraction step also keeps the system efficient. By selecting a manageable set of candidate claims, the
downstream verifier can spend its effort on the statements that are most likely to matter for the final answer.

\section{Verification }

The Verification agent in \path{agents/verification.py} is the analytical heart of the implementation. For each
claim it iterates over the sentences of the candidate papers and computes an evidence score that combines Jaccard
overlap with focus-term and content-term agreement and small type- and source-aware bonuses. Each scored sentence
is then classified as supporting or contradicting the claim, using negation and contradiction-cue analysis so that
a polarity mismatch between claim and sentence is treated as a contradiction.

From the collected evidence the agent selects a verdict. An overgeneralized claim returns insufficient evidence
unless a contradiction pass is explicitly requested; strong, multi-source counter-evidence yields a contradicted
verdict; adequate aligned support yields a supported verdict; and the absence of sufficient overlap yields
insufficient evidence. Confidence is computed by \texttt{\_supported\_confidence} and \texttt{\_contradicted\_confidence}
from the strongest evidence overlap, the count of additional supporting snippets, and the number of independent
corroborating sources, with an explicit penalty when a claim is corroborated only by its own source. All confidence
values are clamped between a floor and a ceiling so the system never reports absolute certainty. The optional
language-model refinement is invoked only after this deterministic assessment is formed, and only over the
evidence already selected, which follows the broader trend of mixing deterministic grounding with optional model
assistance in verification pipelines \cite{kadavath2022language,shinn2023reflexion,asai2024selfrag}.

The result is a verifier that behaves predictably while still leaving room for nuance. Clear evidence leads to a
strong verdict, mixed evidence leads to a more measured conclusion, and uncertainty remains visible rather than
being flattened away.

\section{Critique and Adaptive Re-Verification }

The Critic agent in \path{agents/critic.py} implements both reporting and adaptivity. Its \texttt{critique} method
produces notes for insufficient, low-confidence, single-source, and contradicted claims, including details such as
whether a single-source claim is merely self-corroborated. Its \texttt{identify\_reverification\_targets} method
selects which weak claims warrant another pass and builds a follow-up query from the claim text, its focus terms,
its source title, and a counter-evidence probe when the claim uses absolute or contradiction-prone language.
Overgeneralized claims are excluded, because additional retrieval cannot settle a claim that is too broad. This is
the implementation of the bounded, targeted adaptivity required in Chapter~3 and echoes the critique-driven loops
seen in recent agentic reasoning systems \cite{wu2023autogen,yao2023treeofthoughts,asai2024selfrag}.

This section is especially important in the thesis narrative because it shows that the system does not stop at
flagging weaknesses. It also identifies the specific cases where another evidence pass is likely to help, which is
exactly the kind of controlled adaptivity that makes the framework research-oriented.

\section{Synthesis and Reporting }

The Synthesis agent in \path{agents/synthesis.py} aggregates the assessments into a calibrated summary, optionally
rewritten by the language model under instructions to stay grounded and avoid overclaiming, with a deterministic
fallback. The Report agent in \path{agents/report.py} renders the structured \texttt{FinalReport}, including the
retrieved papers, the orchestration trace, and the full per-claim verification details with evidence snippets and
overlap scores. Because the report is both a structured object and rendered Markdown, the same result can be shown
in the interface, returned by the API, or exported, which is a useful property for evaluation and reproducibility
in retrieval-grounded systems \cite{es2023ragas,min2023factscore,lin2021truthfulqa}.

This implementation directly supports the user-facing experience. The final answer appears as a concise synthesis,
while the verified claims remain available for inspection, which gives the system both readability and
accountability.

\section{The Optional Language-Model Layer }

The language-model abstraction in \path{llm.py} wraps a provider client behind a single interface. It is available
only when an API key is configured; otherwise the agents run on their deterministic logic. The module is written to
be provider-tolerant: a helper detects whether the configured model accepts a reasoning-effort parameter and
includes it only for reasoning-capable models, while models that would reject it are called without it. Structured
outputs are parsed into small typed schemas for the refined verdict and the rewritten summary, and any call failure
is logged and degrades to the deterministic result rather than breaking the pipeline, which reflects the common
practice of using hosted language models as an optional refinement layer rather than as the sole source of truth
\cite{kadavath2022language,shinn2023reflexion,wu2023autogen}.

This design serves three purposes. First, it makes the core pipeline reproducible and offline-capable. Second, it
ensures the model can improve fluency and borderline verdicts but can never bypass the grounding and honesty rules,
which run first. Third, it keeps the system model-agnostic, so a different model or provider can be substituted
through configuration.

That balance is useful in a thesis setting because it keeps the implementation stable and practical while still
leaving room for refinement where language quality matters.

\section{Live Scholarly Retrieval }

Live evidence is implemented in \path{agents/live_retrieval.py}. The agent queries PubMed Central through the NCBI
E-utilities (search, summary, and fetch) and arXiv through its Atom API. Two implementation details are important
for evidence quality. First, PubMed Central searches are sorted by relevance rather than date, because the default
date ordering surfaces recent but off-topic papers. Second, abstracts in the PubMed Central XML are nested inside
structured elements, so the agent gathers all descendant text rather than only the top-level text node, which is
what allows it to recover full abstracts instead of empty placeholders. The agent also derives cleaner search
variants from a natural-language question by removing generic filler words, and it identifies itself to NCBI and
backs off on rate-limit responses, optionally using an API key to raise the request rate.

These details matter because they turn live retrieval into a dependable evidence source rather than a best-effort
search wrapper. In the thesis, that supports the claim that live evidence can be integrated without sacrificing
traceability or control, consistent with modern biomedical and scholarly retrieval practice \cite{gu2021pubmedbert,lee2019biobert,thakur2021beir}.

\section{Configuration and Scope }

Behaviour is centralized in the \texttt{Settings} dataclass in \path{config.py}, which holds the retrieval depth,
the coverage and support thresholds, the confidence floor and ceiling, the source-priority bonuses, the model
configuration, and the keyword lists used by the scope guard. Settings can be constructed directly or built from
environment variables, which is how the interfaces enable live retrieval and the language model when the
appropriate keys are present. The scope guard in \path{scope.py} uses the non-research keyword list to decline
clearly non-research queries while admitting research questions from any academic domain, returning an explicit
out-of-scope message that names the offending topic when possible.

Centralizing these values keeps the implementation easy to tune and easy to describe in the thesis. It also makes
the behaviour of the full system more transparent, because the thresholds and source preferences are defined in one
place rather than scattered across the codebase.

\section{Ingestion of User Papers }

User-provided papers enter the system through \path{ingestion.py}, which parses Markdown, plain-text, JSON, and PDF
inputs into \texttt{PaperDocument} records. JSON is the most reliable format because it carries explicit metadata;
Markdown and text files are parsed for front-matter or a leading abstract paragraph; and PDFs are handled with a
best-effort text extractor. This is the mechanism by which the highest-priority evidence layer --- the user's own
papers --- is brought into the retrieval pool.

This ingestion path is important because it gives the user direct control over provenance. Papers that the user
chooses to upload become part of the evidence environment immediately, which supports the overall thesis goal of
making the system adaptable to the research context at hand.

\section{Baselines as First-Class Clients }

The single-agent and RAG baselines in \path{baselines.py} are implemented as disciplined clients of the same
retrieval and scoring code rather than as separate systems. The single-agent baseline retrieves the top paper and
trusts a direct overlap check, while the RAG baseline retrieves several papers and aggregates simple supporting
overlap without explicit contradiction analysis; both cap their confidence to reflect their weaker evidential
discipline. Implementing the baselines on the shared substrate ensures that the comparison in the evaluation
isolates the effect of orchestration, claim-level verification, and critique, rather than confounding it with
unrelated differences in retrieval, matching the ablation logic used in retrieval-augmented evaluation studies
\cite{petroni2020kilt,thakur2021beir,ragsurvey2024}.

This makes the comparison fair and thesis-appropriate. The baselines are simple enough to highlight what the full
system adds, but they are still grounded in the same retrieval and evidence machinery, which keeps the evaluation
credible.

\section{Interfaces }

The Streamlit application in \path{app.py} exposes the full pipeline interactively: the user uploads papers, selects
a model, asks a question, and sees the executive summary, the retrieved papers with their source labels (uploaded,
corpus, or live API), the orchestration trace, and the per-claim verdicts with confidence and evidence. The FastAPI
service exposes the same capability programmatically. Both interfaces construct settings that enable live retrieval
and the language model from the environment, and both rely entirely on the package's reasoning logic, so the
interface and the studied system are one and the same.

This is a useful design choice for the thesis because it means the same reasoning path can be demonstrated
interactively during a viva and also exercised programmatically in evaluation scripts.

\section{Implementation Principles }

Several principles recur across the codebase:

\begin{enumerate}
\item \textbf{Graceful degradation.} The language model and live retrieval are additive; the system retains full
deterministic functionality when they are unavailable.
\item \textbf{Grounding before generation.} Deterministic evidence scoring and verdict selection run first and
constrain any model refinement.
\item \textbf{Bounded effort.} Retrieval depth, claim counts, and the re-verification loop are bounded to keep the
workflow focused and predictable.
\item \textbf{Modularity.} Retrieval, extraction, verification, critique, synthesis, and reporting are separable
enough to be reused by the baselines and the interfaces.
\item \textbf{Traceability.} Each stage contributes explicit, typed outputs and trace lines that can be inspected
later by the user.
\end{enumerate}

These principles are what make AI-Scientist more than a prompt chain. They turn it into a controlled software
system whose behaviour can be studied systematically.

Together, they explain why the implementation can be presented as a research system rather than only as a
prototype. The code is modular, inspectable, and configured around the same evidence-first design that the thesis
argues for throughout the earlier chapters.

\section{Chapter Conclusion }

This chapter mapped the AI-Scientist architecture to its concrete implementation. The pipeline centers on a typed,
traceable set of agents behind a single Orchestrator; retrieval realizes the layered, source-prioritized evidence
model; verification and confidence are computed deterministically and refined optionally; the Critic drives bounded
adaptive re-verification; and the language model and live retrieval are optional enhancements that never override
the grounding rules. The same package drives the user interface, the API, and the baselines. With the
implementation now grounded in code, the next chapter describes how the system is evaluated.
