\begin{thebibliography}{99}
\bibitem{lewis2020rag}
Lewis, Patrick and Perez, Ethan and Piktus, Aleksandra and Petroni, Fabio and others. Retrieval-Augmented Generation for Knowledge-Intensive NLP. Advances in Neural Information Processing Systems (NeurIPS). 2020.

\bibitem{ragsurvey2024}
Gao, Yunfan and Xiong, Yun and Gao, Xinyu and others. Retrieval-Augmented Generation for Large Language Models: A Survey. arXiv:2312.10997. 2024.

\bibitem{ji2023hallucination}
Ji, Ziwei and Lee, Nayeon and Frieske, Rita and others. Survey of Hallucination in Natural Language Generation. ACM Computing Surveys. 2023.

\bibitem{huang2023hallucinationsurvey}
Huang, Lei and Yu, Weijiang and Ma, Weitao and others. A Survey on Hallucination in Large Language Models: Principles, Taxonomy, Challenges, and Open Questions. arXiv:2311.05232. 2023.

\bibitem{manakul2023selfcheckgpt}
Manakul, Potsawee and Liusie, Adian and Gales, Mark J. F.. SelfCheckGPT. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2023.

\bibitem{min2023factscore}
Min, Sewon and Krishna, Kalpesh and Lyu, Xinxi and others. FActScore. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2023.

\bibitem{wadden2020scifact}
Wadden, David and Lin, Shanchuan and Lo, Kyle and others. Fact or Fiction: Verifying Scientific Claims. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2020.

\bibitem{thorne2018fever}
Thorne, James and Vlachos, Andreas and Christodoulopoulos, Christos and Mittal, Arpit. FEVER. Proceedings of the Conference of the North American Chapter of the Association for Computational Linguistics (NAACL). 2018.

\bibitem{yao2023react}
Yao, Shunyu and Zhao, Jeffrey and Yu, Dian and others. ReAct. International Conference on Learning Representations (ICLR). 2023.

\bibitem{shinn2023reflexion}
Shinn, Noah and Cassano, Federico and Gopinath, Ashwin and Narasimhan, Karthik and Yao, Shunyu. Reflexion. Advances in Neural Information Processing Systems (NeurIPS). 2023.

\bibitem{asai2024selfrag}
Asai, Akari and Wu, Zeqiu and Wang, Yizhong and Sil, Avirup and Hajishirzi, Hannaneh. Self-RAG. International Conference on Learning Representations (ICLR). 2024.

\bibitem{dhuliawala2023cove}
Dhuliawala, Shehzaad and Komeili, Mojtaba and Xu, Jing and others. Chain-of-Verification Reduces Hallucination in Large Language Models. arXiv:2309.11495. 2023.

\bibitem{wu2023autogen}
Wu, Qingyun and Bansal, Gagan and Zhang, Jieyu and others. AutoGen. arXiv:2308.08155. 2023.

\bibitem{li2023camel}
Li, Guohao and Hammoud, Hasan Abed Al Kader and Itani, Hani and Khizbullin, Dmitrii and Ghanem, Bernard. CAMEL. Advances in Neural Information Processing Systems (NeurIPS). 2023.

\bibitem{du2023debate}
Du, Yilun and Li, Shuang and Torralba, Antonio and Tenenbaum, Joshua B. and Mordatch, Igor. Improving Factuality and Reasoning in Language Models through Multiagent Debate. arXiv:2305.14325. 2023.

\bibitem{es2023ragas}
Es, Shahul and James, Jithin and Espinosa-Anke, Luis and Schockaert, Steven. RAGAS. arXiv:2309.15217. 2023.

\bibitem{robertson2009bm25}
Robertson, Stephen and Zaragoza, Hugo. The Probabilistic Relevance Framework: BM25. Foundations and Trends in Information Retrieval. 2009.

\bibitem{ncbi_eutils}
National Center for Biotechnology Information. Entrez Programming Utilities (E-utilities. NCBI/NLM developer documentation. 2023.

\bibitem{arxiv_api}
arXiv. arXiv. Cornell University Library. 2023.

\bibitem{langgraph2024}
LangChain. LangGraph. Open-source framework documentation. 2024.

\bibitem{vaswani2017attention}
Vaswani, Ashish and Shazeer, Noam and Parmar, Niki and others. Attention Is All You Need. Advances in Neural Information Processing Systems (NeurIPS). 2017.

\bibitem{devlin2019bert}
Devlin, Jacob and Chang, Ming-Wei and Lee, Kenton and Toutanova, Kristina. BERT. Proceedings of the Conference of the North American Chapter of the Association for Computational Linguistics (NAACL). 2019.

\bibitem{raffel2020t5}
Raffel, Colin and Shazeer, Noam and Roberts, Adam and others. Exploring the Limits of Transfer Learning with a Unified Text-to-Text Transformer. JMLR. 2020.

\bibitem{petroni2019knowledgebases}
Petroni, Fabio and Rockta. Language Models as Knowledge Bases?. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2019.

\bibitem{guu2020realm}
Guu, Kelvin and Lee, Kenton and Tung, Zora and Pasupat, Panupong and Chang, Ming-Wei. REALM. Proceedings of the International Conference on Machine Learning (ICML). 2020.

\bibitem{karpukhin2020dpr}
Karpukhin, Vladimir and Ou g. Dense Passage Retrieval for Open-Domain Question Answering. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2020.

\bibitem{khattab2020colbert}
Khattab, Omar and Zaharia, Matei. ColBERT. Proceedings of the International Conference on Research and Development in Information Retrieval (SIGIR). 2020.

\bibitem{izacard2021fid}
Izacard, Gautier and Grave, Edouard. Leveraging Passage Retrieval with Generative Models for Open Domain Question Answering. Proceedings of the European Chapter of the Association for Computational Linguistics (EACL). 2021.

\bibitem{izacard2022atlas}
Izacard, Gautier and Lewis, Patrick and Lomeli, Maria and others. Atlas. Proceedings of the International Conference on Machine Learning (ICML). 2022.

\bibitem{thakur2021beir}
Thakur, Nandan and Reimers, Nils and Ru. BEIR. Proceedings of the Conference on Neural Information Processing Systems (NeurIPS). 2021.

\bibitem{rajpurkar2016squad}
Rajpurkar, Pranav and Zhang, Jian and Lopyrev, Konstantin and Liang, Percy. SQuAD. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2016.

\bibitem{joshi2017triviaqa}
Joshi, Mandar and Choi, Eunsol and Weld, Daniel and others. TriviaQA. Proceedings of the Association for Computational Linguistics (ACL). 2017.

\bibitem{yang2018hotpotqa}
Yang, Zhilin and Qi, Peng and Zhang, Saizheng and others. HotpotQA. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2018.

\bibitem{fan2020wikimultihopqa}
Ho, Xuan and others. 2WikiMultiHopQA. Proceedings of the Language Resources and Evaluation Conference (LREC). 2020.

\bibitem{chung2022truthfulqa}
Lin, Stephanie and Hilton, Jacob and Evans, Owain. TruthfulQA: Measuring How Models Mimic Human Falsehoods. Proceedings of the Annual Meeting of the Association for Computational Linguistics (ACL). 2022.

\bibitem{lin2021truthfulqa}
Lin, Stephanie and Hilton, Jacob and Evans, Owain. TruthfulQA: Measuring How Models Mimic Human Falsehoods. arXiv:2109.07958. 2021.

\bibitem{nakano2021webgpt}
Nakano, Reiichiro and Hilton, Jacob and Balaji, Sai and others. WebGPT. arXiv:2112.09332. 2021.

\bibitem{schick2023toolformer}
Schick, Timo and Dwivedi-Yu, Jane and Dess`i. Toolformer: Language Models Can Teach Themselves to Use Tools. Proceedings of the International Conference on Learning Representations (ICLR). 2023.

\bibitem{park2023generativeagents}
Park, Joon Sung and O'Brien, Joseph C. and Cai, Carrie J. and others. Generative Agents: Interactive Simulacra of Human Behavior. arXiv:2304.03442. 2023.

\bibitem{yao2023treeofthoughts}
Yao, Shunyu and Yu, Dian and Zhao, Jeffrey and others. Tree of Thoughts: Deliberate Problem Solving with Large Language Models. arXiv:2305.10601. 2023.

\bibitem{beltagy2019scibert}
Beltagy, Iz and Lo, Kyle and Cohan, Arman. SciBERT. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2019.

\bibitem{lee2019biobert}
Lee, Jinhyuk and Yoon, Wonjin and Kim, Sungdong and others. BioBERT. Bioinformatics. 2020.

\bibitem{gu2021pubmedbert}
Gu, Yuhao and Tinn, Robert and Cheng, Hao and others. Domain-Specific Language Model Pretraining for Biomedical Natural Language Processing. arXiv:2007.15779. 2021.

\bibitem{alsentzer2019clinicalbert}
Alsentzer, Emily and Murphy, John R. and Boag, William and others. Publicly Available Clinical BERT. Proceedings of the Clinical Natural Language Processing Workshop. 2019.

\bibitem{jin2019pubmedqa}
Jin, Qiao and Dhingra, Bhuwan and Liu, Zhengping and others. PubMedQA. Proceedings of the Conference on Empirical Methods in Natural Language Processing (EMNLP). 2019.

\bibitem{pal2022medmcqa}
Pal, Ankit and Umapathi, Logesh Kumar and Sankarasubbu, Malaikannan. MedMCQA. Proceedings of the Conference on Health, Inference, and Learning (CHIL). 2022.

\bibitem{jin2021medqa}
Jin, Di and Pan, Eileen and Oufattole, Nassim and others. What Disease Does This Patient Have? A Large-scale Open Domain Question Answering Dataset from Medical Exams. Applied Sciences 11(14):6421. 2021.

\bibitem{honovich2022true}
Honovich, Or and Aharoni, Roee and Herzig, Jonathan and others. TRUE. Proceedings of the Association for Computational Linguistics (ACL). 2022.

\bibitem{ouyang2022instructgpt}
Ouyang, Long and Wu, Jeffrey and Jiang, Xu and others. Training Language Models to Follow Instructions with Human Feedback. arXiv:2203.02155. 2022.

\bibitem{wei2022cot}
Wei, Jason and Wang, Xuezhi and Schuurmans, Dale and others. Chain-of-Thought Prompting Elicits Reasoning in Large Language Models. arXiv:2201.11903. 2022.

\bibitem{kojima2022zeroshotreasoners}
Kojima, Takeshi and Gu, Shixiang Shane and Reid, Maarten and others. Large Language Models are Zero-Shot Reasoners. arXiv:2205.11916. 2022.

\bibitem{wang2022selfconsistency}
Wang, Xuezhi and Wei, Jason and Schuurmans, Dale and others. Self-Consistency Improves Chain of Thought Reasoning in Language Models. arXiv:2203.11171. 2022.

\bibitem{borgeaud2022retro}
Borgeaud, Sebastian and Mensch, Arthur and Hoffmann, Jordan and others. Improving Language Models by Retrieving from Trillions of Tokens. Proceedings of the International Conference on Machine Learning (ICML). 2022.

\bibitem{zhou2022leasttomost}
Zhou, Denny and others. Least-to-Most Prompting Enables Complex Reasoning in Large Language Models. arXiv:2205.10625. 2022.

\bibitem{khandelwal2020knnlm}
Khandelwal, Urvashi and Lewis, Mike and Jurafsky, Dan and Zettlemoyer, Luke. Generalization through Memorization: Nearest Neighbor Language Models. arXiv:1911.00172. 2020.

\bibitem{lewis2020bart}
Lewis, Mike and Liu, Yinhan and Goyal, Naman and others. BART. arXiv:1910.13461. 2020.

\bibitem{petroni2020kilt}
Petroni, Fabio and Piktus, Aleksandra and Fan, Angela and others. KILT. arXiv:2009.02252. 2020.

\bibitem{kadavath2022language}
Kadavath, Saurav and Conerly, Tom and Askell, Amanda and others. Language Models (Mostly) Know What They Know. arXiv:2207.05221. 2022.

\bibitem{kuhn2023semantic}
Kuhn, Lorenz and Gal, Yarin and Farquhar, Sebastian. Semantic Uncertainty: Linguistic Invariances for Uncertainty Estimation in Natural Language Generation. arXiv:2302.09664. 2023.

\bibitem{wang2024self}
Wang, Yile and others. Self-Knowledge Guided Retrieval Augmentation for Large Language Models. arXiv:2310.05002. 2024.

\end{thebibliography}


