\section{Introduction}

AI-Scientist is a systems thesis, but it is also an evaluation thesis. The contribution depends not only on whether
the software runs, but on whether its behaviour can be examined honestly against meaningful baselines on
representative scientific questions. This chapter therefore explains the evaluation methodology used throughout the
project.

The methodology reflects a deliberate response to the risks identified in the literature review. Grounded question
answering is easy to overclaim and hard to assess cleanly, because a fluent answer can hide whether it is actually
supported by evidence. The central AI-Scientist claim concerns grounding, calibrated confidence, contradiction
handling, abstention, and traceability rather than raw fluency. The evaluation is therefore organized around those
qualities, uses real scholarly evidence rather than synthetic self-confirming text, and compares the multi-agent
system against the two conventional alternatives it is meant to improve upon.

\section{Evaluation Philosophy}

The first methodological principle is that a fluent answer is not automatically a good one. A response counts as
trustworthy only when its conclusions are tied to retrieved evidence, when its confidence reflects the strength and
independence of that evidence, and when its uncertainty is reported rather than hidden. The evaluation therefore
inspects the structure of the system's output --- its per-claim verdicts, confidence, evidence snippets, and
orchestration trace --- not only the surface text of the summary.

The second principle is that the comparison must isolate the effect of the design under study. The single-agent and
RAG baselines are implemented on the same retrieval and scoring substrate as the full system, so any observed
difference reflects orchestration, claim-level verification, and critique rather than incidental differences in
retrieval or tokenization.

The third principle is that honest, cautious behaviour is a success, not a failure. When the evidence for a claim
is thin or contradictory, the correct outcome is an insufficient-evidence or contradicted verdict. The evaluation
explicitly credits the system for abstaining and for surfacing disagreement, rather than rewarding it for always
producing a confident-sounding conclusion.

\section{Baselines}

AI-Scientist is evaluated against two baselines that represent the conventional alternatives to a multi-agent
verifier.

The \textbf{single-agent baseline} models a direct reader: it retrieves the single most relevant paper and trusts a
direct overlap check between the question and that paper's sentences. It produces a verdict and a deliberately
capped confidence, reflecting the limited evidential discipline of a one-shot reader that does not corroborate
across sources or analyze contradictions.

The \textbf{RAG baseline} models a standard retrieval-augmented pipeline: it retrieves several papers and aggregates
simple supporting overlap into a verdict, again with capped confidence. Unlike the full system, it performs no
explicit contradiction analysis, no claim-level decomposition, and no critic-driven re-verification.

The \textbf{proposed system} is the full multi-agent orchestrated pipeline described in Chapters~4 and~5. Holding
the retrieval substrate fixed across all three systems makes the comparison a clean ablation of the contributions
of orchestration, claim-level verification, confidence calibration, and critique.

\section{The Cross-Domain Question Set}

Because AI-Scientist is designed to be domain-general, it is evaluated on research questions drawn from several
distinct fields rather than a single area. The question set spans medicine, computer science and artificial
intelligence, biology, and psychology, with additional probes in physics, chemistry, and the life sciences to
exercise the live-retrieval path. Representative questions include whether metformin improves weight and metabolic
outcomes in type~2 diabetes, whether retrieval-augmented generation reduces hallucination in language models,
whether CRISPR/Cas9 correction is a promising treatment for sickle cell disease, and how social-media use relates
to adolescent depression and affect.

These questions are chosen for three reasons. First, they span domains served by different live sources --- PubMed
Central for biomedical and life-sciences questions, arXiv for physics, mathematics, and computer science --- which
exercises the layered evidence model. Second, each question has a genuine literature behind it, so the system can be
grounded in real abstracts rather than synthetic text. Third, the set deliberately mixes questions with strong,
clear evidence and questions where the evidence is mixed or thin, so the evaluation can observe both confident
support and honest abstention.

\section{Evidence Conditions}

Each question is evaluated under the layered evidence model. Where appropriate, a small set of genuinely relevant
papers is supplied as session uploads to represent the user-provided evidence layer; the curated offline corpus
provides an always-available grounding; and live retrieval from PubMed Central and arXiv supplies current results
at query time. Crucially, the evidence is real: abstracts are the actual text returned by the scholarly sources,
not fabricated, self-confirming passages. This matters because a verifier evaluated only on self-supporting
synthetic text would appear far more confident than one tested against the genuine variability of the literature.

Reporting the source of every retrieved paper --- uploaded, curated corpus, or live API --- is part of the
methodology. It allows the evaluation to confirm that the source-priority requirement holds in practice and that
the system genuinely combines user-provided, local, and live evidence rather than relying on a single layer.

\section{Qualities Assessed}

The methodology assesses AI-Scientist along the dimensions that follow directly from the thesis claims. These
correspond to the evaluation axes set out in the project proposal and to the faithfulness-oriented dimensions
emphasized in the grounded-generation literature.

\textbf{Evidence grounding (faithfulness).} Whether each supported conclusion is actually backed by a retrieved
sentence, inspected through the per-claim evidence snippets. A grounded system should never present a claim as
supported when no evidence sentence aligns with it.

\textbf{Claim verification behaviour.} Whether the verdicts --- supported, contradicted, or insufficient evidence
--- are appropriate to the available evidence, including the correct use of abstention for thin evidence and of
the contradicted verdict when counter-evidence dominates.

\textbf{Confidence calibration.} Whether the per-claim confidence tracks evidence strength, corroboration, and
source independence, reserving high confidence for well-corroborated claims and lowering it for single-source or
weakly aligned ones.

\textbf{Contradiction handling.} Whether the system surfaces disagreement in the literature explicitly rather than
silently choosing the more agreeable side.

\textbf{Hallucination resistance.} Whether the system avoids asserting claims that the evidence does not support, a
property that follows from grounding and abstention together.

\textbf{Reasoning traceability.} Whether the orchestration trace and per-claim evidence make the path to each
conclusion reconstructable by a reader.

\textbf{Domain generality.} Whether the system behaves consistently across domains and correctly declines clearly
non-research queries through the scope guard.

\section{Evaluation Protocol}

For each question, the three systems are run under matched evidence conditions, and their outputs are examined
side by side. The full system's report is inspected for its verdicts, confidence values, evidence attribution,
critique notes, and whether the critic-driven re-verification loop is triggered. The baselines' outputs are
examined for the same question to see what a single-agent reader and a standard RAG pipeline produce on the same
evidence.

The protocol is qualitative and behaviour-oriented by design. Rather than reducing trustworthiness to a single
aggregate score, it presents representative case studies that show how the system grounds, calibrates, and
criticizes its conclusions, and how that behaviour differs from the baselines. This case-study orientation is
appropriate for a verification system whose central claims are about the \emph{structure} and \emph{honesty} of its
reasoning, properties that a single accuracy number would obscure.

\section{Reproducibility}

The evaluation is designed to be reproducible. The deterministic core of the pipeline --- retrieval ranking,
evidence scoring, verdict selection, confidence computation, and critique --- runs without any external model, so
the same question over the same evidence yields the same structured assessment. Live retrieval introduces the
natural variability of the scholarly sources, which is disclosed rather than hidden: a question asked on different
days may surface a slightly different set of current papers. Where the optional language model is used, it refines
fluency and borderline verdicts on top of the deterministic result, so the grounding and confidence behaviour does
not depend on a particular model being available.

\section{Honesty Rules for the Write-Up}

Two practical rules govern the reporting of the evaluation. First, the system is credited for cautious behaviour:
an insufficient-evidence verdict on a genuinely thin question is reported as correct conduct, not as a miss.
Second, confidence is interpreted with its context: a high confidence is meaningful only when accompanied by
independent corroboration, and the write-up reports the corroboration alongside the confidence rather than the
number alone. These rules keep the evaluation aligned with the thesis claim that AI-Scientist values dependable,
evidence-attributed answers over confident but unsupported ones.

\section{Chapter Conclusion}

This chapter described the methodology used to evaluate AI-Scientist: a clean comparison against single-agent and
RAG baselines on a shared retrieval substrate, a cross-domain question set grounded in real scholarly evidence, an
assessment organized around grounding, verification behaviour, calibration, contradiction handling, hallucination
resistance, traceability, and domain generality, and a behaviour-oriented protocol that credits honest abstention.
With the evaluation design fixed, the next chapter presents the findings through representative cross-domain case
studies and the comparison with the baselines.
