\chapter*{Acknowledgment}

It is with a feeling of profound gratitude and regards that I acknowledge the invaluable guidance and encouragement rendered to me by my deeply respected supervisor Dr. Avantika Singh of the Department of Data Science and Artificial Engineering, IIIT Naya Raipur throughout the course of my research work. Without the meticulous care and attention with which she supervised me and constantly reviewed the entire work it would not have been possible to carry out the work to this level. I would like to extend a special thank you to Aakansha Baghel and Yogesh Kumar Sahu for their exceptional support and encouragement. I also express my gratitude to Aditya Pundlik Helode, Dr. Mayank Lovanshi, Mayank Sharma, Nilanjan Mitra, Siddharth Kumar, Tarun Kashyap, Temin Nishad, and Vishnu Prasad Verma for their unwavering support and guidance throughout my thesis, which greatly contributed to the successful completion of my research work.

In addition, I extend my heartfelt gratitude to all the faculty members of the CSE and Data Science departments for their unwavering support. I am also thankful to the lab assistants of the AIDL, Optical Lab, and Big Data Labs for providing me with the opportunity to conduct my research work from the very first semester. 


%I want to expressed my gratitude to all my teacher who have bestowed their knowledge to make me what I am.

%My hearty regards to all my teachers who not only have crafted 

%I would also like to thank my teachers from my school life till now whose has brought me here

%I would also like to thank my teachers - Mr. Nanda Lal Bera, Mr. Sachindra Nath Mathata, Mr. Pabitra Chakroborty, Mr. Souman Show (Poiet), Mr. Sudhir Chandra Rana, Dr. Mrinmoy Hota, Dr. R. Mishra, Mr. L. N. Sahu, Prof. B. N. Biswas, Dr. L. M. Joshie whose thought has brought me here.


%I like to say a big \lq thank you\rq~to all my friends and fellow students for the tremendous support I received from them during my stay at IIT Kharagpur and during the preparation of my thesis. I shall be failing in my duty if I do not name Dr. C. B. Ashesh, Dr. Supriyo Paul, Animesh Goldar, Vamsi Krishna Velidi, Guntupalli Ajay Babu, Dr. Shamik Sarkar, Haradhan Jal \& Chandan Shyamal.
 
%I can ever give enough thanks to my parents, brother, and other family members for their encouragement and support.

%Finally, I am grateful to my wife, Piku, for her patience, devotion, help and encouragement during my stay at IIT Kharagpur and the preparation of this thesis. 

%Most of the work which required computers was done using computers running the GNU/Linux operating system. The software I used was almost entirely from the Open Source movement and I am convinced that without the excellent work of all those GNUians out there, my life would have been more difficult. I wish to thank all those anonymous contributors to the Open Source movement and hope to become part of it by contributing something substantial someday.


%\vspace*{3.50cm}
%\begin{flushright}
%\begin{figure}[htbp]
%  \rightline{\psfig{figure=Maifuz_Sig_2.ps,height=1.50cm}}
%\end{figure}
%{\em (Maifuz Ali)}
%\end{flushright}


\vspace*{1.5cm}
\begin{flushright}
%\begin{figure}[htbp]
%  \rightline{\psfig{figure=/home/maifuzali/LaTex/own/Signature/Maifuz_Sig_2.ps,height=1.750cm}}
%\end{figure}
%\raisebox{0.1cm}[0.0cm][0cm]{
\begin{minipage}{7.0cm}
{\em (Urmila)}
\end{minipage}
%}
\end{flushright}







\vspace*{-1.3cm}
\noindent
\noindent{\ May 22, 2025}\\
 Naya Raipur\\
