\section{Introduction}

The previous chapter established that grounded scientific answering is already a recognized research task, but
that the most defensible remaining gap lies in \emph{verification-first}, \emph{evidence-attributed}, and
\emph{honest} answering. This chapter turns that positioning into a precise problem statement. Rather than treating
AI-Scientist as a generic question-answering agent, the thesis defines it as a constrained verification system: it
observes a research question, gathers layered evidence, decomposes the relevant material into discrete claims,
assigns each claim a grounded verdict and confidence, criticizes the weak claims, and reports the result with a
traceable rationale.

This precision matters for two reasons. First, it prevents overclaiming. A thesis that vaguely states
``AI-Scientist answers research questions'' would blur together retrieval, claim extraction, verification, and
synthesis. Second, a well-defined problem statement makes the evaluation meaningful. The qualities examined later
in the thesis measure not only whether an answer is produced, but whether the answer is grounded, calibrated, and
inspectable.

\section{Task Definition}

AI-Scientist addresses the following task: given a natural-language research question \(q\) from any academic
domain, and an evidence environment consisting of optional user-uploaded papers, a curated offline corpus, and
optional live scholarly sources, produce a structured report that contains a grounded conclusion supported by
traceable evidence. The task is therefore not merely ``generate an answer.'' It is a structured verification
problem over noisy and incomplete evidence.

At a high level, the system input consists of:

\begin{enumerate}
\item the research question or claim \(q\) submitted by the user,
\item an optional set of user-uploaded papers supplied for the session, and
\item an evidence environment: a curated corpus that is always available offline, and optional live retrieval from
PubMed Central and arXiv.
\end{enumerate}

The system output is a structured report containing an executive summary, the list of retrieved papers, a set of
per-claim assessments, a set of critique notes, and an orchestration trace. In the implementation this structure is
explicit: each assessment materializes a claim, a verdict, a confidence, the supporting or contradicting evidence
snippets, and a rationale. Making this structure explicit, rather than leaving it implicit in free-form text, is a
deliberate design choice because it makes the system measurable and auditable.

\section{Verdict Space}

The central abstraction in AI-Scientist is the \emph{verified claim} rather than the free-form answer. Each
extracted claim is assigned exactly one verdict from a bounded set:

\begin{enumerate}
\item \textbf{Supported:} retrieved evidence aligns with the claim, with confidence weighted by evidence strength,
corroboration, and source independence.
\item \textbf{Contradicted:} the available evidence trends against the claim, or strong counter-evidence coexists
with partial support; the disagreement is reported explicitly rather than hidden.
\item \textbf{Insufficient evidence:} the evidence pool does not provide enough overlapping support to settle the
claim; the system abstains instead of overclaiming.
\item \textbf{Out of scope:} the query is not a research question and is declined before analysis begins.
\end{enumerate}

The thesis treats \emph{insufficient evidence} as a first-class, desirable outcome rather than a failure. In a
verification system, declining to assert an unsupported claim is the correct behaviour, and much of the value of
AI-Scientist lies in preferring an honest abstention over a confident fabrication. This is the verification-first
analogue of the abstention principle that recurs throughout the trustworthy-LLM literature.

\section{The Layered Evidence Model}

Because grounding is the foundation of every verdict, the problem definition fixes where evidence may come from
and how the sources are prioritized. AI-Scientist uses three evidence layers:

\begin{enumerate}
\item \textbf{User-uploaded papers:} documents the user supplies for the current session, treated as the most
authoritative source because they reflect the user's intent and provenance.
\item \textbf{Curated corpus:} an offline collection of real paper records that is always available, giving the
system a deterministic grounding even without network access.
\item \textbf{Live scholarly sources:} results retrieved at query time from PubMed Central and arXiv, which keep
the evidence current and broaden coverage beyond the local material.
\end{enumerate}

The layers are not treated as equal. The problem definition imposes a strict source priority --- uploaded papers
outrank curated-corpus papers, which in turn outrank live-API papers --- so that user-provided and locally
trusted evidence is preferred when relevance is comparable. Concretely, a paper's local rank score combines its
relevance to the query with a source-dependent bonus,
\begin{equation}
\mathrm{score}(d, q) \;=\; \min\!\big(1,\; \mathrm{cov}(q, d) + b_{s(d)}\big),
\qquad b_{\mathrm{upload}} > b_{\mathrm{corpus}} > b_{\mathrm{live}},
\end{equation}
where \(s(d)\) is the source of document \(d\) and \(\mathrm{cov}(q,d)\) is the query-coverage relevance defined
below. This makes the source-priority requirement an explicit, inspectable property of retrieval rather than an
emergent accident of ranking.

\section{Relevance and the Coverage Measure}

A recurring subproblem is deciding whether a paper is relevant enough to a query to serve as evidence. A natural
first choice is Jaccard token overlap,
\begin{equation}
\mathrm{ovl}(a,b) \;=\; \frac{|T(a) \cap T(b)|}{|T(a) \cup T(b)|},
\end{equation}
where \(T(\cdot)\) is the set of normalized content tokens of a text. Jaccard overlap is appropriate when comparing
two short texts, such as a claim and a single sentence, and the system uses it for sentence-level evidence scoring.
However, it is a poor signal for ranking whole papers against a short query, because a long abstract inflates the
union \(|T(a) \cup T(b)|\) and drives the score down even when the abstract contains every query term. A genuinely
relevant paper can thus be pushed below a relevance threshold purely because its abstract is long.

To avoid this length bias, AI-Scientist ranks and filters candidate papers with a length-robust query-coverage
measure,
\begin{equation}
\mathrm{cov}(q, d) \;=\; \frac{|T(q) \cap T(d)|}{|T(q)|},
\end{equation}
the fraction of the query's content tokens that appear in the document. A paper is admitted as evidence only if its
coverage exceeds a minimum threshold \(\tau_{\mathrm{cov}}\). This separation --- coverage for paper-level ranking,
Jaccard overlap for sentence-level evidence --- is part of the problem definition because it directly determines
which evidence the verifier ever sees.

\section{Confidence as a First-Class Output}

The problem definition requires that every verdict carry an explicit confidence \(c \in [c_{\min}, c_{\max}]\),
bounded away from both \(0\) and \(1\) so that the system never reports absolute certainty. For a supported claim,
confidence increases with the strength of the best supporting evidence, with the amount of additional
corroborating evidence, and with the number of \emph{independent} sources --- sources other than the paper the
claim was extracted from --- and it is penalized when a claim is corroborated only by its own source paper.
Schematically,
\begin{equation}
c \;=\; \mathrm{clip}\Big(\beta_0 + \beta_1\,\mathrm{ovl}_{\max} + \beta_2\,n_{\mathrm{extra}}
+ \beta_3\,n_{\mathrm{indep}} - \gamma\,\mathbb{1}[\text{self-only}],\; c_{\min},\; c_{\max}\Big),
\end{equation}
where \(\mathrm{ovl}_{\max}\) is the strongest evidence overlap, \(n_{\mathrm{extra}}\) is the count of additional
supporting snippets, \(n_{\mathrm{indep}}\) is the number of independent corroborating sources, and the indicator
penalizes claims supported only by their own source. The exact coefficients are part of the implementation; the
requirement at the problem level is that confidence be a transparent function of evidence strength, corroboration,
and independence, never a free-form number emitted by a model.

\section{Scope: A Domain-General Research Assistant}

AI-Scientist is intentionally domain-general. The system is designed to accept research questions from medicine,
biology, computer science, physics, psychology, economics, and other academic fields, because the verification
machinery --- retrieve, extract claims, check against evidence, criticize, synthesize --- does not depend on the
subject area. Domain generality is therefore a scope decision, not an afterthought.

At the same time, AI-Scientist is not a general chatbot. A scope guard rejects clearly non-research, commercial, or
entertainment queries before any analysis begins, returning an explicit out-of-scope message. The guard is
deliberately permissive about academic subject matter and restrictive only about non-research intent: it admits a
question unless it matches a list of non-research cues such as shopping, product recommendations, or personal
advice. This keeps the system broad across legitimate research domains while preventing it from being misused as a
generic assistant for which its verification design provides no benefit.

\section{Verification and Honesty Requirements}

The core requirement of AI-Scientist is that an answer must never be detached from its evidence. Several concrete
requirements follow from this commitment.

\begin{enumerate}
\item \textbf{Evidence-bound verdicts.} A claim may be marked supported only if at least one retrieved sentence
aligns with it above the support threshold; otherwise it is insufficient or contradicted.
\item \textbf{Explicit contradiction handling.} When supporting and contradicting evidence coexist, the system must
be able to report a contradicted verdict and describe the disagreement, rather than silently choosing the more
agreeable side.
\item \textbf{Abstention over overclaiming.} Overgeneralized claims and claims without adequate overlap must return
insufficient evidence rather than a fabricated conclusion.
\item \textbf{Source independence awareness.} Confidence must account for whether corroboration comes from
independent papers or only from the claim's own source, so that self-corroboration does not masquerade as
consensus.
\item \textbf{Traceability.} Each verdict must retain its evidence snippets and rationale, and the run must record
an orchestration trace describing the steps taken.
\end{enumerate}

These requirements embody the thesis claim that trustworthiness must be engineered explicitly. Because the
verification and confidence logic is encoded as transparent rules over retrieved evidence, the later evaluation can
examine the system's behaviour directly rather than inferring it informally.

\section{Adaptive Reasoning Requirement}

Beyond single-pass verification, the problem definition requires \emph{adaptive} reasoning. The system must be able
to recognize that a first-pass verdict is weak --- because it is insufficient, low-confidence, or single-source ---
and respond by gathering more targeted evidence and re-verifying that specific claim. This requirement is what
distinguishes AI-Scientist from a fixed RAG pipeline: the depth of effort spent on a claim is conditioned on how
uncertain the first pass was, rather than being uniform across all claims.

This adaptivity is bounded. The re-verification loop is targeted at specific weak claims and is constrained so that
the system does not iterate indefinitely. The goal is disciplined additional effort where it is warranted, not
open-ended search.

\section{Non-Goals}

Several tempting claims are explicitly outside the scope of this thesis.

\begin{enumerate}
\item \textbf{Autonomous scientific discovery} is not claimed. AI-Scientist verifies and reports on existing
literature; it does not design experiments or generate novel hypotheses as established findings.
\item \textbf{Full-text deep reading} is not claimed. The system reasons primarily over titles, abstracts, and
uploaded paper records, not over the complete text of every paper.
\item \textbf{Guaranteed factual correctness} is not claimed. A grounded verdict is stronger evidence than an
ungrounded assertion, but it remains bounded by the quality and coverage of the retrieved evidence.
\item \textbf{Exhaustive literature coverage} is not claimed. Live retrieval surfaces a relevant subset of sources,
not the entire body of work on a topic.
\item \textbf{Replacing human judgement} is not the goal. AI-Scientist is designed as an inspectable assistant
whose conclusions a human can audit, not as a final authority.
\end{enumerate}

These non-goals are important because they protect the thesis from being judged against claims it does not need to
make. A well-scoped verification assistant is a stronger contribution than an over-ambitious but indefensible
oracle.

\section{From Problem Statement to Architecture}

The formal problem statement in this chapter now gives a clean lens for reading the rest of the thesis.
AI-Scientist accepts a research question, gathers layered and source-prioritized evidence, decomposes it into
claims, verifies each claim against evidence with explicit confidence, criticizes the weak claims and re-verifies
them adaptively, and reports the result with a traceable rationale. The system is therefore defined not by answer
generation alone, but by a disciplined pipeline that links retrieval, verification, critique, and synthesis.

This framing also clarifies why the architecture in the next chapter looks the way it does. The orchestrated agent
pipeline, the layered evidence model, the confidence function, the critic-driven re-verification loop, and the
scope guard are not independent engineering tricks. They are architectural consequences of the problem definition
established here.

\section{Chapter Conclusion}

This chapter has defined the AI-Scientist task, its verdict space, the layered evidence model, the relevance and
confidence measures, the verification and honesty requirements, the adaptive-reasoning requirement, and the scope
boundaries. These definitions make the later architecture and evaluation measurable: AI-Scientist is judged not
only by whether it can produce an answer, but by whether it grounds, calibrates, criticizes, and traces that answer
under explicit honesty constraints. The next chapter translates these requirements into the system architecture.
