\section{Introduction}

The AI-Scientist problem sits at the intersection of several active research areas rather than belonging to only
one of them. It is related to the study of hallucination in large language models, to retrieval-augmented
generation, to automated fact-checking and scientific claim verification, and to the growing literature on
multi-agent reasoning and self-critique \cite{ji2023hallucination,ragsurvey2024,wadden2020scifact,wu2023autogen}.
A good related-work chapter for AI-Scientist therefore needs to do more than list nearby papers. It must explain
which parts of the prior literature AI-Scientist inherits directly, which capabilities have already been
established by others, and which gap still remains meaningful enough to support a thesis contribution.

As of the time of writing, the field has moved well beyond the point where one can claim novelty simply by saying
that a model retrieves evidence before answering. Retrieval-augmented and self-reflective systems already
demonstrate that grounding and self-critique reduce hallucination. The important question is no longer whether
grounding should be used at all, but rather how scientific answering should be \emph{structured},
\emph{verified}, and \emph{made transparent}. This chapter reviews the most relevant prior work and shows why
AI-Scientist is best positioned not as ``the first grounded research assistant,'' but as an orchestrated,
verification-first, evidence-attributed system built on top of an already established grounding setting.

\section{Hallucination in Large Language Models}

The motivating problem for AI-Scientist is hallucination: the tendency of generative models to produce statements
that are fluent and confident but unsupported by, or contradictory to, the underlying facts. Surveys of the
phenomenon distinguish between intrinsic hallucination, where the output contradicts the provided source, and
extrinsic hallucination, where the output cannot be verified against any available source
\cite{ji2023hallucination,huang2023hallucinationsurvey}. Both are damaging in a scientific setting, but extrinsic
hallucination is especially dangerous because it includes fabricated findings and invented citations that a
non-expert reader cannot easily detect.

This literature matters for AI-Scientist in two ways. First, it establishes that hallucination is not a rare edge
case but a structural property of how current models generate text, which justifies building verification into the
system rather than trusting the base model. Second, it motivates the central design decision of the thesis: to
make the evidence for each statement explicit, so that intrinsic hallucination becomes checkable against the
retrieved sources and extrinsic hallucination is surfaced as ``insufficient evidence'' rather than presented as
fact.

A related strand of work focuses on \emph{detecting} hallucination after generation. SelfCheckGPT, for example,
samples multiple responses and measures their consistency as a black-box signal of factuality, on the intuition
that hallucinated content varies more across samples than grounded content \cite{manakul2023selfcheckgpt}.
AI-Scientist shares the underlying goal of flagging untrustworthy statements, but takes a complementary route:
rather than detecting hallucination by self-consistency alone, it grounds each claim in retrieved evidence and
assigns confidence from the strength and independence of that evidence.

\section{Measuring Factual Precision}

Detecting hallucination is only useful if factuality can be measured at a meaningful granularity. Coarse,
response-level judgments hide the fact that a long answer may be partly correct and partly fabricated. FActScore
addresses this by decomposing a generated passage into atomic facts and scoring each against a knowledge source,
yielding a fine-grained precision measure \cite{min2023factscore}. This decomposition principle is directly
relevant to AI-Scientist, whose Claim Extraction agent performs an analogous step: it breaks the retrieved
material into discrete, individually checkable claims so that verification and confidence can be reported per
claim rather than per answer.

The lesson AI-Scientist takes from this line of work is methodological. A trustworthy research assistant should
not output a single monolithic verdict on a whole answer. It should expose the unit of evidence at the level of
individual claims, because that is the level at which support, contradiction, and uncertainty actually differ.

\section{Retrieval-Augmented Generation}

The closest architectural ancestor of AI-Scientist is retrieval-augmented generation. The original RAG formulation
conditions a generator on passages retrieved from an external corpus, improving performance on knowledge-intensive
tasks and reducing reliance on parametric memory \cite{lewis2020rag}. Subsequent surveys document a rich design
space covering how to index, retrieve, rerank, and fuse evidence, and how to balance retrieval cost against answer
quality \cite{ragsurvey2024}. AI-Scientist inherits the core RAG insight that grounding generation in retrieved
text reduces fabrication, and its Retrieval agent and layered evidence model are squarely within this tradition.

However, standard RAG also sets a baseline that AI-Scientist must improve upon rather than merely reproduce. In a
typical RAG pipeline, retrieval is a silent preprocessing step and the model still emits a single fused answer. The
reader sees neither which passages actually supported each part of the answer nor how confident the system is in
each component. Self-RAG sharpens this picture by training a model to decide when to retrieve and to critique its
own generations with reflection tokens, moving RAG from a fixed pipeline toward an adaptive, self-evaluating one
\cite{asai2024selfrag}. AI-Scientist pursues the same adaptive spirit, but realizes it through an explicit
multi-agent division of labour and an external critic rather than through learned reflection tokens, which keeps
the verification logic inspectable and model-agnostic.

For AI-Scientist, the most useful interpretation of the RAG literature is that grounding is necessary but not
sufficient. Retrieval supplies the raw material for verification; it does not, on its own, tell the user which
claims are well-supported, which rest on a single source, and which are contradicted.

\section{Automated Fact-Checking and Scientific Claim Verification}

A second direct ancestor is automated fact-checking, and in particular scientific claim verification. The FEVER
task framed fact verification as retrieving evidence sentences and labeling a claim as supported, refuted, or
lacking information \cite{thorne2018fever}. SciFact specialized this setting to scientific claims, requiring a
system to find abstracts that support or contradict a claim and to identify the specific rationale sentences
\cite{wadden2020scifact}. This labeling scheme --- supported, contradicted, or insufficient evidence --- maps
almost directly onto the verdict space used by the AI-Scientist Verification agent.

This correspondence is important for positioning. AI-Scientist does not claim to invent the idea of labeling a
scientific claim against retrieved evidence; that task is well established. What the fact-checking literature
generally does not provide is an end-to-end \emph{assistant} that takes a free-form research question, extracts its
own claims from retrieved papers, verifies them, criticizes the weak ones, and synthesizes a confidence-annotated
report for a human user across arbitrary domains. AI-Scientist is best understood as embedding the claim-level
verification paradigm inside a usable, orchestrated research workflow.

The fact-checking tradition also reinforces a key design constraint: the evidence rationale must be explicit. A
verdict without the supporting sentence is not auditable. AI-Scientist therefore attaches, to each verdict, the
specific evidence snippets and their overlap with the claim, mirroring the rationale-selection requirement of
scientific claim verification.

\section{Reasoning, Self-Critique, and Verification Loops}

A third strand of related work concerns how language-model systems reason and check themselves over multiple
steps. ReAct interleaves reasoning traces with actions such as retrieval, improving grounding and interpretability
\cite{yao2023react}. Reflexion adds a verbal self-feedback loop in which an agent reflects on a failed attempt and
retries with that feedback \cite{shinn2023reflexion}. Chain-of-Verification has a model draft an answer, generate
verification questions, answer them independently, and revise, explicitly reducing hallucination through a
self-checking loop \cite{dhuliawala2023cove}.

These systems are conceptually close to the AI-Scientist Critic agent and its re-verification loop. The shared
idea is that a first attempt should not be trusted as final, and that a structured second look can catch errors the
first pass missed. AI-Scientist differs in where it places the critique. Rather than asking one model to critique
itself --- which risks the critique inheriting the same blind spots as the original answer --- it assigns critique
to a separate agent with explicit rules for flagging insufficient, low-confidence, single-source, and contradicted
claims, and it routes the resulting re-verification at the granularity of the individual weak claim. This makes the
loop transparent and targeted rather than global and opaque.

\section{Multi-Agent Frameworks}

The final and most architecturally relevant strand is the recent wave of multi-agent LLM frameworks. AutoGen
provides infrastructure for multiple conversational agents that cooperate and use tools \cite{wu2023autogen};
CAMEL studies role-playing communicative agents \cite{li2023camel}; and multi-agent debate shows that having
several model instances argue and converge can improve factuality and reasoning over a single instance
\cite{du2023debate}. Orchestration libraries such as LangGraph make it practical to express such systems as
explicit state machines over cooperating actors \cite{langgraph2024}.

This environment creates both a risk and an opportunity for AI-Scientist. The risk is that a vague claim such as
``we built a multi-agent system'' will sound derivative, because multi-agent scaffolding is now common. The
opportunity is that most multi-agent demonstrations emphasize open-ended collaboration or debate, whereas
AI-Scientist uses a small, fixed set of specialized roles wired into a deterministic orchestration with a clear
purpose: verify claims against layered evidence and report the result honestly. The contribution is therefore not
the existence of multiple agents, but the disciplined, verification-oriented way they are composed.

The debate literature is particularly instructive. Multi-agent debate improves factuality by forcing disagreement
to surface before consensus \cite{du2023debate}. AI-Scientist obtains a related benefit more cheaply: instead of
several generalist agents debating, a single Critic agent applies explicit weakness criteria and can demand more
evidence, which provides much of the value of adversarial review without the cost and unpredictability of free
debate.

\section{Evaluating Grounded Systems}

Because AI-Scientist makes claims about grounding and trustworthiness, it must be read alongside the literature on
how such systems are evaluated. Frameworks such as RAGAS propose reference-light metrics for retrieval-augmented
systems, including faithfulness (whether the answer is entailed by the retrieved context), answer relevance, and
context relevance \cite{es2023ragas}. These dimensions correspond closely to the qualities AI-Scientist is
designed to optimize: faithfulness maps to evidence grounding, and context relevance maps to the quality of
retrieval ranking.

The evaluation literature also cautions that grounded systems are easy to over-credit. A system can appear faithful
simply by hedging, or appear confident simply by ignoring contradictory evidence. This caution shapes the
evaluation philosophy adopted later in this thesis, which emphasizes whether the system grounds, calibrates, and
exposes its reasoning, rather than only whether it produces an agreeable-sounding answer.

\section{Infrastructure: Scholarly Retrieval Sources}

Finally, AI-Scientist depends on concrete scholarly retrieval infrastructure. Live evidence is drawn from PubMed
Central through the NCBI E-utilities programming interface \cite{ncbi_eutils} and from arXiv through its public API
\cite{arxiv_api}. Classical lexical relevance ideas, exemplified by the BM25 family of ranking functions
\cite{robertson2009bm25}, inform the token-overlap and coverage measures used to rank candidate evidence. These
are not contributions of the thesis, but they are essential building blocks, and the way AI-Scientist combines
them with user uploads and a curated corpus is part of its evidence-layering design.

\section{Implications for AI-Scientist Design}

Taken together, the literature suggests several design consequences for AI-Scientist.

First, grounding is necessary but insufficient. RAG reduces hallucination, but the user still needs per-claim
support, confidence, and contradiction information that a fused answer does not provide.

Second, factuality should be measured at the claim level. FActScore and scientific claim verification both argue
for decomposing answers into atomic, individually checkable units.

Third, verdicts must carry their evidence. The fact-checking tradition shows that a label without a rationale is
not auditable.

Fourth, self-critique helps, but it is most trustworthy when it is structured and external rather than left to the
same model that produced the answer.

Fifth, multi-agent orchestration is valuable not for its novelty but for the discipline it imposes: a fixed set of
specialized roles with a clear verification purpose is easier to reproduce, inspect, and reason about than
free-form agent collaboration.

\section{Chapter Conclusion and Positioning}

The literature reviewed in this chapter leads to a clear and disciplined positioning for AI-Scientist. Prior work
has already demonstrated that grounding reduces hallucination, that factuality can be measured at the claim level,
that scientific claims can be labeled against retrieved abstracts, and that self-critique and multi-agent
collaboration can improve reasoning. At the same time, prior work also makes clear that standard grounded systems
still tend to emit single fused answers without exposing per-claim support, confidence, and disagreement to the
user.

AI-Scientist therefore does not claim novelty in grounding or in the use of multiple agents. Its contribution lies
in composing these established ideas into a \emph{verification-first, evidence-attributed, self-critical research
assistant} in which every conclusion is a checked claim with a calibrated confidence and a traceable rationale.
That framing is supported, not contradicted, by the literature. The next chapter turns from this background to the
formal problem definition, requirements, and scope of the AI-Scientist system.
