\begin{thebibliography}{99}

\bibitem{logsage2025}
LogSage: An LLM-Based Framework for CI/CD Failure Detection and Remediation with Industrial Validation.
ASE 2025 Industry Showcase; arXiv:2506.03691, 2025.

\bibitem{cirepairbench2026}
CI-Repair-Bench: A Repository-Aware Benchmark for Automated Patch Validation via CI Workflows.
arXiv:2604.27148, 2026.

\bibitem{hirebuild2018}
HireBuild: An Automatic Approach to History-Driven Repair of Build Scripts.
Proceedings of the International Conference on Software Engineering, 2018.

\bibitem{deepdelta2019}
DeepDelta: Learning to Repair Compilation Errors.
Proceedings of the ACM Joint European Software Engineering Conference and Symposium on the Foundations of Software Engineering, 2019.

\bibitem{llmaprsurvey2025}
A Survey of LLM-Based Automated Program Repair: Taxonomies, Design Paradigms, and Applications.
arXiv:2506.23749, 2025.

\bibitem{llmaprslr2024}
A Systematic Literature Review on Large Language Models for Automated Program Repair.
arXiv:2405.01466, 2024.

\bibitem{swebench2023}
SWE-bench: Can Language Models Resolve Real-World GitHub Issues?
Repository-level software engineering benchmark, 2023.

\bibitem{agentless2025}
Agentless: Demystifying LLM-Based Software Engineering Agents.
Structured localization and repair pipeline for SWE-bench-style tasks, 2025.

\bibitem{ghalogs2025}
GHALogs: A Large-Scale Dataset of GitHub Actions Workflow Runs and Logs.
MSR 2025 dataset; Zenodo record 10154920, 2025.

\bibitem{gitbugactions2024}
GitBug-Actions: Building Reproducible Benchmarks for GitHub Actions Bugs.
Proceedings of the International Conference on Software Engineering, 2024.

\bibitem{ghataxonomy2025}
Why Do GitHub Actions Workflows Fail?
ACM Transactions on Software Engineering and Methodology, 2025.

\bibitem{ghaflaky2026}
Understanding and Detecting Flaky Builds in GitHub Actions.
arXiv:2602.02307, 2026.

\bibitem{aprtop2025}
Is the Cure Still Worse Than the Disease? Test Overfitting by Large Language Models in Automated Program Repair.
arXiv:2511.16858, 2025.

\bibitem{compass2026}
ComPass: Contrastive Patch Correctness Assessment.
arXiv:2602.07561, 2026.

\bibitem{abstentionSurvey2026}
Know Your Limits: A Survey of Abstention in Large Language Models.
Survey of abstention methods for LLM systems, 2026.

\bibitem{abstentionBench2025}
AbstentionBench: Reasoning About When Language Models Should Abstain.
Benchmark for abstention behavior in language models, 2025.

\bibitem{calibrationCodeRevision2026}
Fine-Grained Confidence Calibration of Large Language Models in Automated Code Revision.
arXiv:2604.06723, 2026.

\bibitem{sgagent2026}
SGAgent: An Agentic Framework for Software Engineering Localization and Repair.
arXiv:2602.23647, 2026.

\end{thebibliography}
